Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-01-11 _paper.rmd
updated chocolate case study but it will need more work

The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels. Each of which was labeled as being either milk or dark chocolates. We can see if a manufacturer accurately portrays the chocolate with this data. We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value. We should note that not all chocolates consist wholly of chocolate. The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable. A random forest model is fit classifying the type of chocolate. We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85% cocoa.

@@

Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of `sugar`. We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label. While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are. This feels related a measurement theory problem.

>>>>

The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark. Dark chocolates are typically considered healthier and often can be marketed at a higher price. It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest. <!--We should note that chocolates often have other ingredients will decrease the predictive power of the model nutritional explanatory variable. Hot sure that this is correct because these were all plain chocolate not caramel chocolate or with nuts. --> A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22). It is compared with a correctly classified dark chocolate.

@@

Figure \@ref(fig:casechocolates) similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of `sugar`. We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label. While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.

XXX Attribution projection shows difference between milk and dark is mostly fiber. Is this one, considered similar to the small group of (predicted) milk chocolates that have high fiber and low sugar? there has to be another variable that is generating this little bubble.

XXX Need to re-do. Not sure that sugar is the right variable to modify. Teh difference between milk and dark is typically fiber, and sugars. So maybe this is a low fiber dark chocolate? It is funny that in whatever view we look at it there is no doubt that this looks like a dark chocolate! Fiber has the most contribution

### 2022-01-11 _paper.tex
updated chocolate case study but it will need more work

The chocolates dataset consists of 88 observations of ten nutritional measurements from their labels. Each of which was labeled as being either milk or dark chocolates. We can see if a manufacturer accurately portrays the chocolate with this data. We are curious to see if chocolates that nutritionally look like milk chocolates are labeled as dark chocolates, which may hold a higher market value. We should note that not all chocolates consist wholly of chocolate. The addition of other ingredients will decrease the predictive power of the model nutritional explanatory variable. A random forest model is fit classifying the type of chocolate. We selected a chocolate labeled dark, through predicted to be milk chocolate compared with a chocolate labeled 85\% cocoa.

@@

Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of \texttt{sugar}. We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label. While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are. This feels related a measurement theory problem.

>>>>

The chocolates dataset consists of 88 observations of ten nutritional measurements determined from their labels, and labeled as being either milk or dark. Dark chocolates are typically considered healthier and often can be marketed at a higher price. It could be interesting to examine the nutritional properties of any dark chocolates that have been misclassified as milk, to determine if a manufacturer is being honest. A random forest model is used, and a chocolate labeled dark but predicted to be milk is chosen as the observation of interest (chocolate 22). It is compared with a correctly classified dark chocolate.

@@

Figure \ref{fig:casechocolates} similarly shows that attribution-space is more separable than data-space. Interestingly, the class imbalance that we suspected was not observed; there are only six chocolates labeled as dark and predicted as milk, while eight of the inverse case. We look at one extreme instance and vary the contribution of \texttt{sugar}. We see notice 2 instances that have clearly well within the milk chocolate dispite being observed as dark chocolate on their product label. While conjector, there are plausible reasons that a manufacturer has incentives to cut corners and label their products different than what they are.

XXX Attribution projection shows difference between milk and dark is mostly fiber. Is this one, considered similar to the small group of (predicted) milk chocolates that have high fiber and low sugar? there has to be another variable that is generating this little bubble.

XXX Need to re-do. Not sure that sugar is the right variable to modify. Teh difference between milk and dark is typically fiber, and sugars. So maybe this is a low fiber dark chocolate? It is funny that in whatever view we look at it there is no doubt that this looks like a dark chocolate! Fiber has the most contribution
